Reintroduction of resistant frogs facilitates landscape-scale recovery in the presence of a lethal fungal disease

### 2024-04-21 doc/manuscript/translocation_bioarxiv.tex
Create final files for submission

To determine whether MYL frog populations show genomic patterns
consistent with an evolutionary response to Bd, we compared frog exomes
(i.e., coding region of a genome) between populations with contrasting
histories of Bd exposure. Specifically, we compared frog genomes sampled
in 4 populations that have not yet experienced a Bd-caused epizootic
(``naive'') \citep{zhou2015} versus in 5 populations that experienced a
Bd epizootic during the past several decades and have since recovered to
varying degrees (``recovering''; Figure~\ref{fig-allelefrequencies})
\citep{knapp2016, vredenburg2010}. Bd-exposure histories of the 9 study
populations are based on 10-20 years of VES and Bd surveillance using
skin swabbing \citep[e.g.,][]{knapp2016, zhou2015, wilber2022}. Naive
populations are characterized by large numbers of adults (i.e, typically
1000s), Bd prevalence that is generally 0\% except during occasional Bd
failed invasions \citep[during which Bd loads remain very
low,][]{wilber2022}, and no history of Bd epizootics since we first
surveyed these populations in the late 1990s and early 2000s
\citep{zhou2015}. In contrast, recovering populations exist in an
enzootic state \citep{briggs2010}, characterized by smaller numbers of
adults (generally \textless{} 500), high Bd prevalence \citep[often
\textgreater{} 80\%,][]{knapp2011}, and, in adults, moderate Bd loads
that are typically well below the level expected to cause mortality
\citep{vredenburg2010}. Naive and recovering populations can be
identified unambiguously using these differences in Bd prevalence and
load. Finally, there is no potential for frog dispersal between the 9
study populations due to intervening distances and topography, as well
as the presence of introduced (predatory) fish and fish-induced habitat
fragmentation.

It is important to note that our genomic analysis is necessarily limited
by the small number of naive populations remaining on the landscape
(nearly all of which we sampled as part of the current study) and their
clumped distribution (Figure~\ref{fig-allelefrequencies}). These
characteristics of naive populations precluded a replicated analyses
among pairs of geographically proximate naive and recovering
populations. Instead, to obtain adequate power, we conducted a pooled
analysis across population types to identify genomic regions showing
patterns consistent with selection. Our comparative approach cannot
provide a direct assessment of the strength or timing of natural
selection, but does allow us to identify specific regions of the genome
that merit further study for their potential contribution to frog
resistance in recovering populations. (See \textbf{Supporting
Information - Frog evolution in response to Bd - Study design} for
additional details regarding the study design.)

@@

Results from the individual variant and splined window analyses show
that recovering populations show patterns consistent with selection on
multiple regions of the genome. The analysis of individual variants
identified 11 ``outlier'' SNPs (i.e., showing significantly different
allele frequencies between naive versus recovering populations) from 7
distinct genes across 4 contigs (Figure S6B, C
SI). One of the 7 identified genes (LOC108802036) does not have an
associated annotation. For the outlier SNPs, frequency differences
between the naive and recovering populations ranged from 0.41 to 0.86.
Most of these SNPs showed frequency differences in only a subset of the
sampled populations (Figure~\ref{fig-allelefrequencies}A, B), but the
SNP in the RIN3 gene showed consistent differences in frequencies across
all populations (Figure~\ref{fig-allelefrequencies}C). This is
suggestive of parallel selection at this locus across multiple
populations. The other 6 outlier variants showed less consistent
frequency differences across the study populations, but for these we
still found statistically significant differences in 2 of the 3 genetic
clusters (containing populations 5--9;
Figure~\ref{fig-allelefrequencies}, Figure S6A). Therefore, although some outlier variant associations have a more
limited geographic extent than RIN3, they still describe results that
are consistent with parallel evolutionary changes following Bd exposure.

The splined window analysis identified 33 outlier regions for
\(\pi_{diff}\) and 58 outlier regions for \emph{F\textsubscript{ST}}
(Figure~\ref{fig-spline-manhattan}A, B). Of these, 9 regions were
outliers for both metrics (``shared regions'') and 2 of these shared
regions also contained one or more of the outlier SNPs described above.
A total of 35 annotated genes were found in the 9 shared regions. Given
this large number of genes, here we focus on those with the strongest
signal of selection and/or immune-related functions. The largest
\(\pi_{diff}\), suggestive of directional selection, occurred in a 163kb
region on Contig19, 12.9Mb upstream of the RIN3 outlier SNP
(Figure~\ref{fig-spline-manhattan}C). This region contains
approximately 500 SNPs and one annotated gene called
``interferon-induced very large GTPase 1-like'' (GVINP1). Additionally,
a shared outlier region on Contig1 contained two complement factor genes
(C6 and C7). Interestingly, this region had a large negative
\(\pi_{diff}\) value, consistent with balancing selection. Finally, one
shared outlier region on Contig8 contained one outlier SNP (TCF19) and
was within 360kb of another outlier SNP (VARS)
(Figure~\ref{fig-spline-manhattan}D, Figure S7).
This region (854kb from the beginning of the outlier window to the VARS
SNP) contained a total of 8 annotated genes. In \emph{Xenopus}, five of
these genes occur in the extended major histocompatibility complex (MHC)
Class I region (FLOT1, TUBB, MDC1, CCHCR1, TCF19) and three occur in the
extended MHC Class III region (HSP70, LSM2, VARS) \citep{ohta2006}.
Therefore, this region is part of the extended MHC Class I and III
complex and shows synteny with other amphibian genomes.

Although the joint processes of Bd-caused population declines and
selection in response to Bd exposure could affect genetic diversity of
recovering populations, we found no consistent differences in
individual-level heterozygosity or population-level \(\pi\) between
naive and recovering populations (see \textbf{Supporting Information -
Frog evolution in response to Bd - Genetic diversity} for details).
Thus, despite patterns consistent with localized selection in particular
regions of the genome, we did not find evidence for reduced genetic
diversity across the genome in recovering populations. In addition, no
gene ontology (GO) biological functions, molecular functions, or
cellular processes were over-represented in either the outlier variants
or the 35 genes located in the overlapping \emph{F\textsubscript{ST}}
and \(\pi_{diff}\) splined windows (see \textbf{Supporting Information -
Frog evolution in response to Bd} \textbf{- GO analysis} for details).

@@

Disease-induced population declines are decimating global biodiversity
\citep{daszak2000}, but broadly-applicable strategies to recover
affected species are generally lacking \citep[e.g.,][]{garner2016}.
Here, we tested the possibility that populations of resistant
individuals from naturally recovering populations can be used to
reestablish extirpated populations of the endangered MYL frog in the
presence of a highly virulent fungal pathogen (Bd). Our results indicate
(i) the capacity of reintroduced populations to become established and
eventually recover despite ongoing disease, (ii) that 50\% of the
recovering populations are likely to persist over a 50-year period,
(iii) that there are genomic differences between naive and recovering
MYL frog populations, consistent with evolutionary change in frogs
following Bd exposure, and (iv) that some of the genomic differences are
in regions that contain genes related to disease resistance.
Collectively, these results (Figure~\ref{fig-recovery-model}) provide a
rare example of amphibian recovery in the presence of Bd, and have
important implications for the conservation and recovery of amphibians
and other taxa worldwide that are endangered by escalating impacts from
emerging infectious diseases. In light of the generally low success rate
of amphibian reintroduction efforts \citep{dodd2005}, our success in
reestablishing MYL frog populations via translocation of resistant
individuals is striking, and even more so given that MYL frogs were
driven to near-extinction by Bd.

@@

In our rapidly changing world, evolution is likely to play in important
role in facilitating the resilience of wildlife populations. Whether the
documented disease resistance in MYL frogs and concurrent recovery of
decimated populations provides an airtight example of evolutionary
rescue will likely always be uncertain (given that we can never have a
perfect understanding of the past). Regardless, we provide an example
from the wild that suggests that evolution might produce individuals
that harbor adaptive alleles and allow population recovery in a novel
(i.e., Bd-positive) environment, and show conclusively that individuals
from these recovering populations can be used to reestablish extirpated
populations and expand the scale of natural recovery
(Figure~\ref{fig-recovery-model}). We expect that similar species
recovery actions will be an essential tool in wildlife conservation in
an era of accelerating global change.

@@

For the 24 translocations we conducted, we identified donor populations
from which adult frogs (\(\geq\) 40 mm snout-vent length) could be
collected using several years of VES and skin swab collections
\citep{knapp2016}, and results from population genetic analyses
\citep{poorten2017}. The populations that we selected contained hundreds
of \emph{R. sierrae} adults and thousands of tadpoles. These relatively
high abundances were the result of recent increases following previous
Bd-caused declines \citep{knapp2016}. As is typical for recovering MYL
frog populations, Bd prevalence in the donor populations was high
(0.69--0.96) and Bd load (median log\textsubscript{10}(load) =
3.06--3.78 ITS copies) was two or more orders of magnitude below the
level at which frog mortality is expected (log\textsubscript{10}(load)
\(\approx\) 5.78 copies)\citep{vredenburg2010, joseph2018}. Recipient
sites to which frogs were translocated were chosen based on previous
\emph{R. sierrae} presence (determined from VES and/or museum records)
or characteristics that suggested high quality habitat for this species
\citep{knapp2005}. At the beginning of this study, we had a relatively
limited understanding of the factors that affect habitat quality. In
subsequent years, we improved our site selection process by
incorporating new information about important habitat features, in
particular, overwinter habitats such as submerged boulders and
overhanging banks. \emph{R. sierrae} were absent from all recipient
sites prior to the first translocation.

@@

Raw reads were filtered for adapters and contaminants using fastp
\citep{chen2018} and aligned to the \emph{Rana muscosa} genome (NCBI
SRA: SRS6533475 \citep{hon2020}), with repetitive elements masked using
bwa (``mem'' mode) \citep{li2013}. Exact PCR duplicates were marked
using Picard. Variants were then called following GATK best practices
(v.4.2.0.0 \citep{van2020}). Briefly, raw variants were called for each
sample using HaplotypeCaller and combined using CombineGVCFs. Next,
genotypes were jointly called using GenotypeGVCFs. Variants were then
hard filtered using gatk VariantFiltration using the following
parameters to remove low-quality sites: QD \textless{} 2.0, FS
\textgreater{} 40.0, SOR \textgreater{} 3.0, MQ \textless{} 50.0,
MQRankSum \textgreater{} 3.0, MQRankSum \textless{} -3.0, ReadPosRankSum
\textgreater{} 3.0, ReadPosRankSum \textless{} -3.0. This initial filter
resulted in 1,595,206 variant sites across 53 individuals.

We then further filtered our dataset at the individual and variant
level. First, we trimmed our variants to only include those with minor
allele frequency \textgreater{} 0.03, a maximum depth of 250 and minimum
depth of 5, a minimum genotype quality of 20, and a maximum missing
proportion of 0.5. This filter resulted in 427,038 sites, of which
353,172 were SNPs and 73,866 were INDELS. Finally, we trimmed samples
with an average depth across filtered sites \textless{} 7x (n = 3). Our
final dataset included 50 samples, 23 from naive and 27 from recovering
populations, with an average depth of 16.7x (range = 7.4x -- 26.1x).

@@

In the splined window analysis, we identified outlier regions using
\emph{F\textsubscript{ST}} and differences in nucleotide diversity
(\(\pi_{diff}\)) between naive and recovering populations. First, we
calculated per-site \emph{F\textsubscript{ST}} between the naive and
recovering individuals for all bi-allelic SNPs in the 30 largest contigs
(98\% of all SNPs) using VCFtools \citep{danecek2011}. Next, we
calculated per-site nucleotide diversity \(\pi\) separately for
individuals from the naive and recovering populations using VCFtools,
then calculated \(\pi_{diff}\) for each population
(\(\pi_{diff} = \pi_{naive} - \pi_{recovering}\)). We concatenated the
values for \emph{F\textsubscript{ST}} and \(\pi_{diff}\) in order of
size-sorted chromosome number and adjusted the SNP position based on the
relative position in the genome (for more efficient data processing and
to better contextualize the strength of the outlier signals in different
regions of the genome). We then used the GenWin R package
\citep{beissinger2015} to conduct a splined discrete window analysis for
\emph{F\textsubscript{ST}} and \(\pi_{diff}\). This method calculates
where non-overlapping window boundaries should occur by identifying
inflection points in the spline fitted to \emph{F\textsubscript{ST}} and
\(\pi_{diff}\) values along the genome, therefore balancing false
positive and false negative results that occur using other window-based
methods \citep{beissinger2015}. This method also calculates a
W-statistic allowing for outlier identification. We identified outliers
as those with a W-statistic greater than 4 standard deviations above the
mean for \emph{F\textsubscript{ST}} or above/below the mean for
\(\pi_{diff}\). These standard deviations represent strict criteria to
select only the top \textasciitilde{} 0.3\% of windows. Shared outliers
were then identified as those that were outliers in both analyses,
meaning that they showed (i) high differentiation between naive and
recovering populations, and (ii) differential patterns of nucleotide
diversity in the same region. Finally, we extracted gene transcripts
mapped within each region and retrieved annotation for that region using
BLAST (\textbf{Supporting Information}: Datasets S3, S4).

@@

\acknow{We thank the following for important contributions to this study: E.
Hegeman and A. Lindauer (project and data management, field and
laboratory work); A. Barbella and K. Rose-(laboratory work); C. Kamoroff
and numerous summer technicians (field work); staff at Sequoia-Kings
Canyon and Yosemite National Parks, Inyo and Sierra National Forests,
California Department of Fish and Wildlife, U.S. Fish and Wildlife
Service, University of California-Santa Barbara Institutional Animal
Care and Use Committee, and Sierra Nevada Aquatic Research Laboratory
(research permits, logistical support, and/or field assistance). This
project was supported by grants from the National Park Service (to
R.A.K.), Yosemite Conservancy (to R.A.K.), and National Science
Foundation (EF-0723563, to C. Briggs; DEB-1557190, to C. Briggs;
DEB-2133401, to M.Q.W; and DBI-2120084, to C. Richards-Zawacki).}

@@

{\centering \includegraphics[width=\textwidth]{figures/conceptual_model.png}

@@

{\centering \includegraphics[width=0.5\textwidth]{figures/translocation_survival_bysiteid.png}

@@

{\centering \includegraphics[width=\textwidth]{figures/cond_effects_plot.png}

@@

{\centering \includegraphics[width=\textwidth]{figures/pop_viability_figures_for_manuscript.jpg}

@@

{\centering \includegraphics[width=\textwidth]{figures/allele_maps.png}

@@

\caption{\label{fig-allelefrequencies}Genomic patterns consistent with
selection on individual variants in recovering MYL frog populations at
the landscape scale (based on a linear mixed model analysis). For each
of the 9 naive and recovering MYL frog populations (indicated by
numbered points), adjacent pie charts show allele frequencies for the 11
outlier SNPs from 7 distinct genes: (A) LOC108802036, (B) TCF19, VARS,
MAP1S, TYRP1, and SRFBP1, and (C) RIN3. Charts are superimposed on a map
of the Sierra Nevada study area, with Yosemite, Kings Canyon, and
Sequoia National Parks (from north to south) shown in dark gray, and the
range boundary for MYL frogs shown in light gray. The inset map locates
the study area and range boundary in California. Sites 1 and 4 (site id
= 72996 and 70567, respectively) were also used as sources of frogs in
the frog population recovery study.}

@@

{\centering \includegraphics[width=\textwidth]{figures/splinewindow_manhattan.png}

@@

\caption{\label{fig-spline-manhattan}Patterns consistent with selection
on genomic regions in recovering MYL frog populations based on a splined
window analysis. Note that several key candidate genes (RIN3, TCF19,
VARS) are outliers in both the individual variant analysis
Figure~\ref{fig-allelefrequencies} and splined window analysis (this
figure). Manhattan plots show outlier regions for the difference in (A)
nucleotide diversity \(\pi_{diff}\) and (B) \emph{F\textsubscript{ST}}.
In (A), outlier regions are shown above the upper red dashed line and
below the lower red dashed line. In (B), outlier regions are shown above
the single dashed red line. Outlier regions for either \(\pi_{diff}\) or
\emph{F\textsubscript{ST}} are shown in blue and outlier regions for
both \(\pi_{diff}\) and \emph{F\textsubscript{ST}} are shown in red. (C)
Magnified Contig19 from (A) showing two adjacent outlier regions for
\(\pi_{diff}\) 12.9Mb upstream of the RIN3 outlier SNP (indicated with a
dashed vertical blue line). (D) Magnified Contig8 from from (B) showing
the \emph{F\textsubscript{ST}} outlier region that includes the outlier
SNPs TCF19 and VARS. This region of the genome contains 8 annotated
genes known to occur in the extended MHC Class I and III regions.}

>>>>

To determine whether MYL frog populations show genomic patterns
consistent with an evolutionary response to Bd, we compared frog exomes
(i.e., coding region of a genome) between populations with contrasting
histories of Bd exposure. Specifically, we compared frog genomes sampled
in 4 populations that have not yet experienced a Bd-caused epizootic
(``naive'') \citep{zhou2015} versus in 5 populations that experienced a
Bd epizootic during the past several decades and have since recovered to
varying degrees (``recovering''; Figure~\ref{fig-allelefrequencies})
\citep{knapp2016, vredenburg2010}. Bd-exposure histories of the 9 study
populations are based on 10-20 years of VES and Bd surveillance using
skin swabbing \citep[e.g.,][]{knapp2016, zhou2015, wilber2022}. Naive
populations are characterized by large numbers of adults (i.e, typically
1000s), Bd prevalence that is generally 0\% except during occasional Bd
failed invasions \citep[during which Bd loads remain very
low,][]{wilber2022}, and no history of Bd epizootics since we first
surveyed these populations in the late 1990s and early 2000s
\citep{zhou2015}. In contrast, recovering populations exist in an
enzootic state \citep{briggs2010}, characterized by smaller numbers of
adults (generally \textless{} 500), high Bd prevalence \citep[often
\textgreater{} 80\%,][]{knapp2011}, and, in adults, moderate Bd loads
that are typically well below the level expected to cause mortality
\citep{vredenburg2010}. Naive and recovering populations can be
identified unambiguously using these differences in Bd prevalence and
load. There is no potential for frog dispersal between the 9 study
populations or any other populations due to intervening distances and
topography, as well as the presence of introduced (predatory) fish and
fish-induced habitat fragmentation.

It is important to note that our genomic analysis is necessarily limited
by the small number of naive populations remaining on the landscape
(nearly all of which we sampled as part of the current study) and their
clumped distribution (Figure~\ref{fig-allelefrequencies}). These
characteristics of naive populations precluded a replicated analysis
among pairs of geographically proximate naive and recovering
populations. Instead, to obtain adequate power, we conducted a pooled
analysis across population types to identify genomic regions showing
patterns consistent with selection. Our comparative approach cannot
provide a direct assessment of the strength or timing of natural
selection, but does allow us to identify specific regions of the genome
that merit further study for their potential contribution to frog
resistance in recovering populations. (See \textbf{Supporting
Information - Frog genome evolution - Study design} for additional
details.)

@@

Results from the individual variant and splined window analyses show
that recovering populations show patterns consistent with selection on
multiple regions of the genome. The analysis of individual variants
identified 11 ``outlier'' SNPs (i.e., showing significantly different
allele frequencies between naive versus recovering populations) from 7
distinct genes across 4 contigs (Figure S6B, C). One of the 7 identified genes (LOC108802036) does not have an
associated annotation. For the outlier SNPs, frequency differences
between the naive and recovering populations ranged from 0.41 to 0.86.
Most of these SNPs showed frequency differences in only a subset of the
sampled populations (Figure~\ref{fig-allelefrequencies}A, B), but the
SNP in the RIN3 gene showed consistent differences in frequencies across
all populations (Figure~\ref{fig-allelefrequencies}C). This is
suggestive of parallel selection at this locus across multiple
populations. The other 6 outlier variants showed less consistent
frequency differences across the study populations, but for these we
still found statistically significant differences in 2 of the 3 genetic
clusters (containing populations 5--9;
Figure~\ref{fig-allelefrequencies}, Figure S6A). Therefore, although some outlier variant associations have a more
limited geographic extent than RIN3, they still describe results that
are consistent with parallel evolutionary changes.

The splined window analysis identified 33 outlier regions for
\(\pi_{diff}\) and 58 outlier regions for \emph{F\textsubscript{ST}}
(Figure~\ref{fig-spline-manhattan}A, B). Of these, 9 regions were
outliers for both metrics (``shared regions'') and 2 of these shared
regions also contained one or more of the outlier SNPs described above.
A total of 35 annotated genes were found in the 9 shared regions. Given
this large number of genes, here we focus on those with the strongest
signal of selection and/or immune-related functions. The largest
\(\pi_{diff}\), suggestive of directional selection, occurred in a 163kb
region on Contig19, 12.9Mb upstream of the RIN3 outlier SNP
(Figure~\ref{fig-spline-manhattan}C). This region contains
approximately 500 SNPs and one annotated gene called
``interferon-induced very large GTPase 1-like'' (GVINP1). Additionally,
a shared outlier region on Contig1 contained two complement factor genes
(C6 and C7). Interestingly, this region had a large negative
\(\pi_{diff}\) value, consistent with balancing selection. Finally, one
shared outlier region on Contig8 contained one outlier SNP (TCF19) and
was within 360kb of another outlier SNP (VARS)
(Figure~\ref{fig-spline-manhattan}D, Figure S7).
This region (854kb from the beginning of the outlier region to the VARS
SNP) contained a total of 8 annotated genes. In \emph{Xenopus}, five of
these genes occur in the extended major histocompatibility complex (MHC)
Class I region (FLOT1, TUBB, MDC1, CCHCR1, TCF19) and three occur in the
extended MHC Class III region (HSP70, LSM2, VARS) \citep{ohta2006}.
Therefore, this region is part of the extended MHC Class I and III
complex and shows synteny with other amphibian genomes. To provide
details specifically on possible selection in the Yosemite populations
described in \textbf{Results - Frog population recovery}, \(\pi\) values
for the 9 shared outlier regions are shown in
Figure S8 for each of the four Yosemite
populations; two of these populations provided frogs for 20 of the 24
frog translocations.

Although the joint processes of Bd-caused population declines and
selection in response to Bd exposure could affect genetic diversity of
recovering populations, we found no consistent differences in
individual-level heterozygosity or population-level \(\pi\) between
naive and recovering populations (see \textbf{Supporting Information -
Frog genomic evolution - Genetic diversity} for details). Thus, despite
patterns consistent with localized selection in particular regions of
the genome, we did not find evidence for reduced genetic diversity
across the genome in recovering populations. In addition, no gene
ontology (GO) biological functions, molecular functions, or cellular
processes were over-represented in either the outlier variants or the 35
genes located in the overlapping \emph{F\textsubscript{ST}} and
\(\pi_{diff}\) splined windows (see \textbf{Supporting Information -
Frog genomic evolution} \textbf{- GO analysis} for details).

@@

Disease-induced population declines are decimating global biodiversity
\citep{daszak2000}, but broadly-applicable strategies to recover
affected species are generally lacking \citep[e.g.,][]{garner2016}.
Here, we tested the possibility that populations of resistant
individuals from naturally recovering populations can be used to
reestablish extirpated populations of the endangered MYL frog in the
presence of a highly virulent pathogen (Bd). Our results indicate (i)
the capacity of reintroduced populations to become established and
eventually recover despite ongoing disease, (ii) that 50\% of the
reintroduced populations are likely to persist over a 50-year period,
(iii) that there are genomic differences between naive and recovering
MYL frog populations, consistent with evolutionary change in frogs
following Bd exposure, and (iv) that some of the genomic differences are
in regions that contain genes related to disease resistance.
Collectively, these results (Figure~\ref{fig-recovery-model}) provide a
rare example of amphibian recovery in the presence of Bd, and have
important implications for the conservation and recovery of amphibians
and other taxa worldwide that are endangered by escalating impacts from
emerging infectious diseases. In light of the generally low success rate
of amphibian reintroduction efforts \citep{dodd2005}, our success in
reestablishing MYL frog populations via reintroduction of resistant
individuals is striking, and even more so given that MYL frogs were
driven to near-extinction by Bd.

@@

In our rapidly changing world, evolution is likely to play in important
role in facilitating the resilience of wildlife populations. We provide
an example from the wild that suggests that evolution might produce
individuals that harbor adaptive alleles and allow population recovery
in a novel (i.e., Bd-positive) environment, and show conclusively that
individuals from these recovering populations can be used to reestablish
extirpated populations and expand the scale of natural recovery
(Figure~\ref{fig-recovery-model}). We expect that similar species
recovery actions will be an essential tool in wildlife conservation in
an era of accelerating global change.

@@

For the 24 translocations we conducted, we identified donor populations
from which adult frogs (\(\geq\) 40 mm snout-vent length) could be
collected using several years of VES and skin swab collections
\citep{knapp2016}, and results from population genetic analyses
\citep{poorten2017}. The populations that we selected contained hundreds
of \emph{R. sierrae} adults and thousands of tadpoles. These relatively
high abundances were the result of recent increases following previous
Bd-caused declines \citep{knapp2016}. As is typical for recovering MYL
frog populations, Bd prevalence in the donor populations was high
(0.69--0.96) and Bd load (median log\textsubscript{10}(load) =
3.06--3.78 ITS copies) was two or more orders of magnitude below the
level at which frog mortality is expected (log\textsubscript{10}(load)
\(\approx\) 5.78 copies)\citep{vredenburg2010, joseph2018}. Recipient
sites to which frogs were translocated were chosen based on previous
\emph{R. sierrae} presence (determined from VES and/or museum records)
or characteristics that suggested high quality habitat for this species
\citep{knapp2005}. At the beginning of this study, we had a relatively
limited understanding of the factors that affect habitat quality. In
subsequent years, we improved our site selection process by
incorporating new information about important habitat features, in
particular, overwintering habitats such as submerged boulders and
overhanging banks. \emph{R. sierrae} were absent from all recipient
sites prior to the first translocation.

@@

Raw reads were filtered for adapters and contaminants using fastp
\citep{chen2018} and aligned to the \emph{R. muscosa} genome (NCBI SRA:
SRS6533475 \citep{hon2020}), with repetitive elements masked using bwa
(``mem'' mode) \citep{li2013}. Exact PCR duplicates were marked using
Picard. Variants were then called following GATK best practices
(v.4.2.0.0 \citep{van2020}). Briefly, raw variants were called for each
sample using HaplotypeCaller and combined using CombineGVCFs. Next,
genotypes were jointly called using GenotypeGVCFs. Variants were then
hard filtered using gatk VariantFiltration using the following
parameters to remove low-quality sites: QD \textless{} 2.0, FS
\textgreater{} 40.0, SOR \textgreater{} 3.0, MQ \textless{} 50.0,
MQRankSum \textgreater{} 3.0, MQRankSum \textless{} -3.0, ReadPosRankSum
\textgreater{} 3.0, ReadPosRankSum \textless{} -3.0. This initial filter
resulted in 1,595,206 variant sites across 53 individuals.

We then further filtered our dataset at the individual and variant
level. First, we trimmed our variants to only include those with minor
allele frequency \textgreater{} 0.03, a maximum depth of 250 and minimum
depth of 5, a minimum genotype quality of 20, and a maximum missing
proportion of 0.5. This filter resulted in 427,038 sites, of which
353,172 were SNPs and 73,866 were INDELS. Finally, we trimmed samples
with an average depth across filtered sites \textless{} 7x (n = 3). Our
final dataset included 50 samples, 23 from naive populations and 27 from
recovering populations, with an average depth of 16.7x (range = 7.4x --
26.1x).

@@

In the splined window analysis, we identified outlier regions using
\emph{F\textsubscript{ST}} and differences in nucleotide diversity
(\(\pi_{diff}\)) between naive and recovering populations. First, we
calculated per-site \emph{F\textsubscript{ST}} between the naive and
recovering individuals for all bi-allelic SNPs in the 30 largest contigs
(98\% of all SNPs) using VCFtools \citep{danecek2011}. Next, we
calculated per-site nucleotide diversity \(\pi\) separately for
individuals from the naive and recovering populations using VCFtools,
then calculated \(\pi_{diff}\) for each population
(\(\pi_{diff} = \pi_{naive} - \pi_{recovering}\)). We concatenated the
values for \emph{F\textsubscript{ST}} and \(\pi_{diff}\) in order of
size-sorted chromosome number and adjusted the SNP position based on the
relative position in the genome (for more efficient data processing and
to better contextualize the strength of the outlier signals in different
regions of the genome). We then used the GenWin R package
\citep{beissinger2015} to conduct a splined discrete window analysis for
\emph{F\textsubscript{ST}} and \(\pi_{diff}\). This method calculates
where non-overlapping window boundaries should occur by identifying
inflection points in the spline fitted to \emph{F\textsubscript{ST}} and
\(\pi_{diff}\) values along the genome, therefore balancing false
positive and false negative results that occur using other window-based
methods \citep{beissinger2015}. This method also calculates a
W-statistic allowing for outlier identification. We identified outliers
as those with a W-statistic greater than 4 standard deviations above the
mean for \emph{F\textsubscript{ST}} or above/below the mean for
\(\pi_{diff}\). These standard deviations represent strict criteria to
select only the top approximately 0.3\% of windows. Shared outliers were
then identified as those that were outliers in both analyses, meaning
that they showed (i) high differentiation between naive and recovering
populations, and (ii) differential patterns of nucleotide diversity in
the same region. Finally, we extracted gene transcripts mapped within
each region and retrieved annotation for that region using BLAST
(\textbf{Supporting Information}: Datasets S3, S4).

@@

\acknow{We thank the following for important contributions to this study: E.
Hegeman and A. Lindauer (project and data management, field and
laboratory work); A. Barbella and K. Rose (laboratory work); numerous
summer technicians (field work); staff at Sequoia-Kings Canyon and
Yosemite National Parks, Inyo and Sierra National Forests, California
Department of Fish and Wildlife, U.S. Fish and Wildlife Service,
University of California-Santa Barbara Institutional Animal Care and Use
Committee, and Sierra Nevada Aquatic Research Laboratory (research
permits, logistical support, and/or field assistance). We also thank the
editor and three anonymous reviewers whose input helped us improve the
manuscript. This project was supported by grants from the National Park
Service (to R.A.K.), Yosemite Conservancy (to R.A.K.), and National
Science Foundation (EF-0723563, to C. Briggs; DEB-1557190, to C. Briggs;
DEB-2133401, to M.Q.W; and DBI-2120084, to C. Richards-Zawacki).

\hypertarget{references}{%
\subsubsection{References}\label{references}}}

@@

{\centering \includegraphics[width=0.95\textwidth]{figures/conceptual_model.png}

@@

{\centering \includegraphics[width=0.60\textwidth]{figures/translocation_survival_bysiteid.png}

@@

{\centering \includegraphics[width=0.80\textwidth]{figures/cond_effects_plot.png}

@@

{\centering \includegraphics[width=0.95\textwidth]{figures/pop_viability_figures_for_manuscript.jpg}

@@

{\centering \includegraphics[width=0.95\textwidth]{figures/allele_maps.png}

@@

\caption{\label{fig-allelefrequencies}Genomic patterns consistent with
selection on individual variants in recovering MYL frog populations at
the landscape scale (based on a linear mixed model analysis). For each
of the 9 naive and recovering MYL frog populations (indicated by
numbered points), adjacent pie charts show allele frequencies for the 11
outlier SNPs from 7 distinct genes: (A) LOC108802036, (B) TCF19, VARS,
MAP1S, TYRP1, and SRFBP1, and (C) RIN3. Charts are superimposed on a map
of the Sierra Nevada study area, with Yosemite, Kings Canyon, and
Sequoia National Parks (from north to south) shown in dark gray, and the
range boundary for MYL frogs shown in light gray. The inset map locates
the study area and range boundary in California. Sites 1 and 4 (site id
= 72996 and 70567, respectively) were also used as sources of frogs for
reintroductions described in \textbf{Results - Frog population
recovery}.}

@@

{\centering \includegraphics[width=0.95\textwidth]{figures/splinewindow_manhattan.png}

@@

\caption{\label{fig-spline-manhattan}Results from the splined window
analysis showing patterns consistent with selection on genomic regions
in recovering MYL frog populations. Note that several key candidate
genes (RIN3, TCF19, VARS) are outliers in both the individual variant
analysis (Figure~\ref{fig-allelefrequencies}) and splined window
analysis (this figure). Manhattan plots show outlier regions for the
difference in (A) nucleotide diversity \(\pi_{diff}\) and (B)
\emph{F\textsubscript{ST}}. In (A), outlier regions are shown above the
upper red dashed line and below the lower red dashed line. In (B),
outlier regions are shown above the single dashed red line. Outlier
regions for either \(\pi_{diff}\) or \emph{F\textsubscript{ST}} are
shown in blue and outlier regions for both \(\pi_{diff}\) and
\emph{F\textsubscript{ST}} are shown in red. (C) Magnified Contig19 from
(A) showing two adjacent outlier regions for \(\pi_{diff}\) 12.9Mb
upstream of the RIN3 outlier SNP (indicated with a dashed vertical blue
line). (D) Magnified Contig8 from from (B) showing the
\emph{F\textsubscript{ST}} outlier region that includes the outlier SNPs
TCF19 and VARS. This region of the genome contains 8 annotated genes
known to occur in the extended MHC Class I and III regions.}
